Yield Mapping

### 2016-02-10 Rmd/yield-predictors.rmd
Updated on DSSAT and empirical modeling side: 1. Created DSSAT-yields.Rmd, for running DSSAT for both maize and soybean. 2. Results of these go into yield-predictors and empirical-yields to create modeled crop yields (using a GAM). 3. zmd-gsod-stations.Rmd creates the station points used in the analysis.

```{r, eval = FALSE}
library(agroEcoTradeoff)
library(rCropMod)
library(matrixStats)
library(doMC)

p_root <- lmisc::set_base_path()
p_crop <- "~/Dropbox/data/agricultural/zambia/dssat/"
p_cropmod <- "~/Dropbox/projects/rCropMod/external/ext_data/"
p_soil <- "~/DSSAT45/SOIL/"
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_clim <- fp(p_root, "external/input_devel/climate")
p_climo <- "~/Dropbox/data/climate/rainfall/zambia/precip/"
p_yield <- fp(p_root, "external/input_devel/yield")
p_pgf <- "~/Dropbox/data/climate/afdm/pgf_bc/zambia/"

@@

# Yield data selection points
# GSOD points
gsod <- fread(fp(p_clim, "gsod.txt"), stringsAsFactors = FALSE)
setnames(gsod, names(gsod), c("nms", "x", "y", "n"))
gsod <- gsod[, 1:3, with = FALSE]
gsod[, nms := paste0("g", nms)]
# plot(idgr)
# points(zamgsod)  # not representative enough

# ZMD rainfall stations
nms <- c('CHIPAT01', 'CHIPEP01', 'CHOMA001', 'ISOKA001', 'KABOMP01', 'KABWE001',
         'KAFIRO01', 'KAFUE001', 'KALABO01', 'KAOMA001', 'KASAMA01', 'KASEMP01',          'KAWAMB01', 'LIVING01', 'LUNDAZ01', 'LUSAKA01', 'LUSAKA02', 'LUSITU01',          'MAGOYE01', 'MANSA001', 'MBALA001', 'MFUWE001', 'MISAMF01', 'MKUSHI01',          'MONGU002', 'MPIKA001') 
x <- c(32.58, 28.38, 27.07, 32.63, 24.20, 28.47, 28.12, 27.92, 22.70, 24.80,
       31.13, 25.85, 29.08, 25.82, 33.20, 28.32, 28.45, 29.37, 27.60, 28.85, 
       31.33, 31.93, 31.25, 29.33, 23.15, 31.43)  
y <- c(-13.55, -17.33, -16.83, -10.12, -13.60, -14.45, -12.60, -15.77, -14.95,
       -14.80, -10.22, -13.53, -9.80, -17.82, -12.28, -15.42, -15.32, -16.30,
       -16.00, -11.10, -8.85, -13.27, -10.10, -14.00, -15.25, -11.90)
zmd <- cbind.data.frame(nms = as.character(nms), x, y, 
                        stringsAsFactors = FALSE)
write.csv(zmd, file = fp(p_climo, "zmd-stations.csv"))

# bring together GSOD and ZMD stations
zmd_gsod <- rbind(zmd, gsod)
coordinates(zmd_gsod) <- c("x", "y")

# Select those intersecting Zambia
gint <- rgeos::gIntersects(prov, zmd_gsod, byid = TRUE)
sel <- unname(which(apply(gint, 1, function(x) sum(which(x))) > 0))
zmet <- zmd_gsod[sel, ]
# str(zmet)

# buffer the ZMD stations by ~20 km to see if any GSOD stations fall within that
zbuffer <- rgeos::gBuffer(zmet[1:26, ], byid = TRUE, width = 0.3)
zbuffer <- spChFIDs(zbuffer, as.character(1:nrow(zbuffer)))
gint <- rgeos::gIntersects(zbuffer, zmet[27:nrow(zmet), ], byid = TRUE)
sel <- unname(which(apply(gint, 1, function(x) sum(which(x))) > 0))
desel <- c(27:nrow(zmet))[sel]
# plot(zmet[-desel, ], pch = 16, cex = 0.5)
# points(zmet[desel, ], pch = 16, cex = 0.5, col = "red")
# plot(prov, add = TRUE)
zmetf <- zmet[-desel, ]

zmetsamps <- extract(idgr, zmetf, cellnumbers = TRUE)
zmetsamps <- cbind(xyFromCell(idgr, zmetsamps[, "cells"]), zmetsamps)

zmet_dt <- data.table(data.frame(zmetsamps))[, .(x, y)]
setkeyv(zmet_dt, c("x", "y"))
setkeyv(IDr, c("x", "y"))
IDr_ss <- IDr[zmet_dt]  # pick up IDs, soils, wth variables
setkeyv(IDr_ss, c("WTH", "Zone"))
# IDr_ss[, plot(x, y)]
# points(zmet_dt, pch = 16, cex = 0.5)
# tst <- extract(idgr, zmet_dt)
# all(tst[order(tst[, 1]), 2] == zmetsamps[order(zmetsamps[, 4]), 5])

@@

# Full DSSAT yield grids
```{r, eval = FALSE}
crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
fylds <- lapply(crops, function(X) {  # X <- "maize"
  fnms <- dir(full_path(p_crop, X), pattern = X)
  fl <- unlist(lapply(c("S", "R"), function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  # idt <- copy(IDr)
  # key(idt)
  idt <- copy(ID)
  idt <- subset(unique(idt[, .(x, y, WTH, Zone, SharePct)]))
  setkeyv(idt, c("WTH", "Zone"))
  for(y in 1:length(fl)) { # y <- 1
    # setcolorder(idt, c(4, 7:8, 5:6, 1:3))
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
    setkeyv(cr, c("WTH", "Zone"))

@@

rdt <- fylds$maize[, c(1:2, grep("YS", names(fylds$maize))), with = FALSE]
mzs <- dt_to_raster(rdt, raster()@crs)
rdt <- fylds$maize[, c(1:2, grep("YR", names(fylds$maize))), with = FALSE]
mzr <- dt_to_raster(rdt, raster()@crs)
rdt <- fylds$soybean[, c(1:2, grep("YS", names(fylds$soybean))), with = FALSE]
soys <- dt_to_raster(rdt, raster()@crs)
rdt <- fylds$soybean[, c(1:2, grep("YR", names(fylds$soybean))), with = FALSE]
soyr <- dt_to_raster(rdt, raster()@crs)

# par(mar = rep(0, 4))
# plot(calc(mzs, mean))
# plot(zmetf, pch = 16, cex = 0.5, add = TRUE)

@@

# Yield data for those points
mannms <- c("S", "R")
ylds <- lapply(crops, function(X) {
  # X <- crops[1]
  fnms <- dir(full_path(p_crop, X), pattern = X)
  fl <- unlist(lapply(mannms, function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  # fl <- fl[1:3]
  crml <- lapply(1:length(fl), function(y) { # y <- 1
    idt <- copy(IDr_ss)
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
    setkeyv(cr, c("WTH", "Zone", "SOL"))

    # merge IDs with subset location
    # crm <- cr[idt]
    crm <- merge(cr, idt)
    # crm[, SharePct := NULL]
    nms <- names(crm)[c(7:8, 6, 1:2, 9, 3, 5, 4)]
    setcolorder(crm, nms)
    mid <- 1:nrow(crm)
    crm <- cbind(mid, crm)
    keys <- c("mid", "x", "y", "ID", "WTH", "Zone", "zoneid", "SOL", "YR")
    setkeyv(crm, keys)
    # setcolorder(crm, c(nms[grep("HWAH", nms, invert = TRUE)], "HWAH"))
    newnm <- gsub(".*._||\\.csv", "", fl[y])
    setnames(crm, "HWAH", newnm)
    crm
  })
  yldmerge <- Reduce(mymerge, crml)
  ynms <- names(yldmerge)
  # yldred <- data.table(round(do.call(cbind, lapply(mannms, function(x) {
    # yldmerge[, rowMeans(.SD), .SDcols = grep(paste0(x, "[0-9]"), ynms)]
  # }))))
  # calculate 
  for(i in mannms) {
    onms <- list(paste0("YLD", i), grep(paste0(i, "[0-9]"), ynms))
    yldmerge[, c(onms[[1]]) := round(rowMeans(.SD)), .SDcols = onms[[2]]]
    yldmerge[, c(onms[[2]]) := NULL]
  }

  # Internal check on yield means
  if(length(crml) == 8) crind <- list("S" = 1:4, "R" = 5:8, "I" = 9:12) 
  if(length(crml) == 6) crind <- list("S" = 1:3, "R" = 4:6, "I" = 7:9) 
  for(ii in sample(1:nrow(yldmerge), size = 20)) {
    for(jj in 1:length(mannms)) {  # jj = 2
      a <- yldmerge[ii, ]
      b <- sapply(crind[[jj]], function(x) {
        crml[[x]][mid == a$mid, ][[ncol(crml[[x]])]]
      })
      if(a[[paste0("YLD", mannms[jj])]] != round(mean(b))) {
        print(paste("Yield mismatch at", ii, mannms[jj]))
      }
    }
  }

@@

@@

## Soil

Note: Using an updated version of `read_dssat_prof.R` from the rCropMod library. It needs to be rebuilt into that library.  Here it is being sourced before runnign the next section. 
```{r, eval = FALSE}
pts <- IDr_ss[, .(x, y)]
# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
coordinates(pts) <- ~x+y
soilid <- extract(soilgrid, pts)  # grid id corresponding to point
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profdat <- read_dssat_prof(solfile = full_path(p_soil, "WI.SOL"), 
                           profiles = profs$SoilProfile) 
solvars <- unique(gsub("[[:digit:]]", "", names(profdat)))
lnms <- names(profdat)

# max depth in profile
mdepth <- rowMaxs(as.matrix(profdat[, .SD, .SDcols = grep("slb", lnms)]), 
                  na.rm = TRUE)

# use depths as weights to calculate mean values of soil parameters to 1 m
# but need to fill to stop at 1 m, and also correct for soils that don't get to 
# 1 m.  Assume values for deepest horizon where deepest < 1 m are the values to 
# 1 m. Truncate horizon that spans 1 m at 1 m, makes this deepest horizon in 
# calc. Horizon depths become weights for weighted mean. 
tst <- profdat[, .SD , .SDcols = grep("slb", lnms)]
dtest <- function(x) {
  o <- x
  o[is.na(o)] <- -10000
  if(all(o < 100)) {
    o[which.max(o)] <- 100
  } else if(any(o > 100)) {
    o[which(o > 100)[1]] <- 100
  }
  o[o < 0] <- NA
  o
}
depvec <- t(apply(as.matrix(tst), 1, dtest))  # fix depths

@@

# calculate mean values to 1 m depth  
solvarsr <- solvars[solvars %in% 
                     c("sbdm", "sloc", "slcl", "slsi", "slhw")]
solvals <- do.call(cbind, lapply(solvarsr, function(i) {  # i <- annvars[2]
  mat <- as.matrix(profdat[, .SD , .SDcols = grep(i, lnms)])
  oval <- sapply(1:nrow(mat), function(j) {  # j <- 6
    ind <- which(!is.na(depvec[j, ]))
    matv <- mat[j, ind]
    matv <- infill(x = matv, ind)
    weighted.mean(matv, depvec[j, ind], na.rm = TRUE)
  })
  oval
}))
colnames(solvals) <- solvarsr

# finalize
sol1m <- cbind("pro" = profdat[, pro], data.table(solvals))
```

@@

# variabes to derive: MAP, N raindays, rainfall depth, n rain days, 
# n dry days in Jan and Feb, mean max temp, N days > 30, annual mean temp, 
# GDD > 30, average daily SRAD
# set up indices for selecting dates
ystrt <- which(mos == 7 & dys == 15) 
ystrt <- ystrt[-length(ystrt)]
yend <- which(mos == 7 & dys == 14)[-1]
# set up seasonal index
sind <- rep(0, length(dts))  # whole year spanning growing season
for(i in 1:length(ystrt)) sind[ystrt[i]:yend[i]] <- i
sindu <- unique(sind)[-1]
janind <- rep(0, length(dts))  # january of each year
for(i in sindu) janind[which(sind == i & mos == 1)] <- i
febind <- rep(0, length(dts))  # feb of each year
for(i in sindu) febind[which(sind == i & mos == 2)] <- i
ndjfind <- rep(0, length(dts))  # djf of each growing season
for(i in sindu) ndjfind[which(sind == i & mos %in% c(1, 2, 11, 12))] <- i
octind <- rep(0, length(dts))  # feb of each year
for(i in sindu) octind[which(sind == i & mos == 10)] <- i

@@

@@

# growing degree day and killing degree days
kdd <- stack(kdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], kdd))
writeRaster(kdd, filename = fp(p_clim, "kdd.tif"))
gdd <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gdd830))
writeRaster(gdd, filename = fp(p_clim, "gdd.tif"))

# tmax greater than 30, Jan and Feb
nkilljan <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(janind == sindu2[i])
  calc(tmax[[ind]] - 273.15, ftmax30)
})
writeRaster(nkilljan, filename = fp(p_clim, "nkilljan.tif"))
nkillfeb <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(febind == sindu2[i])
  calc(tmax[[ind]] - 273.15, ftmax30)
})
writeRaster(nkillfeb, filename = fp(p_clim, "nkillfeb.tif"))

# mean tmax in NDJF
tmax_gs <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  calc(tmax[[ind]] - 273.15, mean)
})
writeRaster(tmax_gs, filename = fp(p_clim, "tmax_gs.tif"))

# mean tmax whole year
tmax_ano <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  calc(tmax[[ind]] - 273.15, mean)
})
writeRaster(tmax_ano, filename = fp(p_clim, "tmax_ano.tif"))

>>>>

```{r, eval = FALSE}
library(agroEcoTradeoff)
library(gdalUtils)
library(rCropMod)
library(matrixStats)
library(doMC)

p_root <- lmisc::set_base_path()
p_crop <- "~/Dropbox/data/agricultural/zambia/dssat/"
p_cropmod <- "~/Dropbox/projects/rCropMod/external/ext_data/"
p_soil <- "~/DSSAT45/SOIL/"
p_basedat <- paste0("/Users/lestes/Dropbox/projects/agroEcoTradeoff/external/",
                    "base_data/")
p_clim <- fp(p_root, "external/input_devel/climate")
p_climo <- "~/Dropbox/data/climate/rainfall/zambia/precip/"
p_yield <- fp(p_root, "external/input_devel/yield")
p_pgf <- "~/Dropbox/data/climate/afdm/pgf_bc/zambia/"
p_sgrids <- "~/Dropbox/data/soil/"

@@

# Yield data selection points
load(fp(p_yield, "empirical/stations.rda"))
zmetsamps <- extract(idgr, zmetf, cellnumbers = TRUE)
zmetxy <- data.frame(zmetf)[, c("x", "y")]
# xFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$x
# yFromCell(idgr, zmetsamps[, "cells"]) - zmetxy$y
# zmetsamps <- cbind(xyFromCell(idgr, zmetsamps[, "cells"]), zmetsamps)
zmetsamps <- cbind(zmetxy, zmetsamps)  # harmonize with DSSAT zmetfxy

# zmet_dt <- data.table(data.frame(zmetsamps))[, .(x, y)]
zmet_dt <- data.table(data.frame(zmetsamps))
# setkeyv(zmet_dt, c("x", "y"))
setkeyv(zmet_dt, c("ID", "zoneid"))

# setkeyv(IDr, c("x", "y"))
setkeyv(IDr, c("ID", "zoneid"))
IDr_ss <- merge(IDr, zmet_dt)
# IDr_ss <- IDr[zmet_dt]  # pick up IDs, soils, wth variables
IDr_ss[, c("x.x", "y.x", "cells") := NULL]
setnames(IDr_ss, c("x.y", "y.y"), c("x", "y"))
setkeyv(IDr_ss, c("WTH", "Zone"))

# Collect soils up here also 
pts <- IDr_ss
# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
coordinates(pts) <- ~x+y
soilid <- extract(soilgrid, pts)  # grid id corresponding to point
IDr_ss[, soilid := soilid]
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profs[, SharePct := NULL]
IDr_ss <- merge(IDr_ss, profs, by.x = "soilid", by.y = "CELL5M")
IDr_ss[, soilid := NULL]
setnames(IDr_ss, "SoilProfile", "SOL")
setkeyv(IDr_ss, c("Zone", "WTH", "SOL"))

# IDr_ss[, plot(x, y)]
# points(zmet_dt, pch = 16, cex = 0.5)
# tst <- extract(idgr, zmet_dt)
# all(tst[order(tst[, 1]), 2] == zmetsamps[order(zmetsamps[, 4]), 5])

@@

# Full DSSAT yield grids
```{r, eval = FALSE}
# crops <- dir(p_crop, pattern = "maize|soybean", full.names = FALSE)
crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
fylds <- lapply(crops, function(X) {  # X <- "maize"
  fnms <- dir(full_path(p_crop, X), pattern = X)
  fl <- unlist(lapply(c("S", "R"), function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  # idt <- copy(IDr)
  # key(idt)
  idt <- copy(ID)
  idt <- subset(unique(idt[, .(x, y, WTH, Zone, SharePct)]))
  setkeyv(idt, c("WTH", "Zone"))
  for(y in 1:length(fl)) { # y <- 1
    # setcolorder(idt, c(4, 7:8, 5:6, 1:3))
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
    setkeyv(cr, c("WTH", "Zone"))

@@

rdt <- fylds$maize[, c(1:2, grep("YS", names(fylds$maize))), with = FALSE]
mzs <- dt_to_raster(rdt, raster()@crs)
rdt <- fylds$maize[, c(1:2, grep("YR", names(fylds$maize))), with = FALSE]
mzr <- dt_to_raster(rdt, raster()@crs)

# par(mar = rep(0, 4))
plot(calc(mzs, max))
plot(calc(mzr, mean))
# plot(calc(soyr, mean))
# plot(zmetf, pch = 16, cex = 0.5, add = TRUE)

@@

crops <- dir(p_crop, pattern = "maize", full.names = FALSE)  # only maize now
# Yield data for those points
mannms <- c("S", "R")
ylds <- lapply(crops, function(X) {
  # X <- crops[1]
  fnms <- dir(full_path(p_crop, X), pattern = X)
  fl <- unlist(lapply(mannms, function(x) {
    fnms[grep(paste0(paste0(x, "0"), "|", paste0(x, "9")), fnms)]
  }))  # reorder to read subsistence first
  # fl <- fl[1:3]
  crml <- lapply(1:length(fl), function(y) { # y <- 1
    idt <- copy(IDr_ss)
    print(paste("Processing", fl[y]))
    cr <- fread(input = full_path(p_crop, paste0(X, "/", fl[y])))
    cr[, "YR" := RUNNO - min(RUNNO) + 1, by = .(TNAM, Zones)]  # unique year  
    setnames(cr, c("SOIL_ID...", "WSTA....", "Zones"), c("SOL", "WTH", "Zone"))
    cr[, WTH := substr(WTH, 1, 4)]
    cr[, c("V1", "RUNNO", "TRNO", "TNAM") := NULL]
    setkeyv(cr, c("Zone", "WTH", "SOL"))

    # merge IDs with subset location
    # crm <- cr[idt]
    crm <- merge(cr, idt)
    # crm[, SharePct := NULL]
    nms <- names(crm)[c(7:8, 6, 1:2, 9, 3, 5, 4)]
    setcolorder(crm, nms)
    mid <- 1:nrow(crm)
    crm <- cbind(mid, crm)
    keys <- c("mid", "x", "y", "ID", "WTH", "Zone", "zoneid", "SOL", "YR")
    setkeyv(crm, keys)
    # setcolorder(crm, c(nms[grep("HWAH", nms, invert = TRUE)], "HWAH"))
    newnm <- gsub(".*._||\\.csv", "", fl[y])
    setnames(crm, "HWAH", newnm)
    crm
  })
  yldmerge <- Reduce(mymerge, crml)
  ynms <- names(yldmerge)
  # yldred <- data.table(round(do.call(cbind, lapply(mannms, function(x) {
    # yldmerge[, rowMeans(.SD), .SDcols = grep(paste0(x, "[0-9]"), ynms)]
  # }))))
  # calculate 
  for(i in mannms) {
    onms <- list(paste0("YLD", i), grep(paste0(i, "[0-9]"), ynms))
    # yldmerge[, c(onms[[1]]) := round(rowMeans(.SD)), .SDcols = onms[[2]]]
    # Selecting rowMax here
    yldmerge[, c(onms[[1]]) := apply(.SD, 1, max), .SDcols = onms[[2]]]
    yldmerge[, c(onms[[2]]) := NULL]
  }

  # Internal check on yield means
  if(length(crml) == 8) crind <- list("S" = 1:4, "R" = 5:8, "I" = 9:12) 
  if(length(crml) == 6) crind <- list("S" = 1:3, "R" = 4:6, "I" = 7:9) 
  for(ii in sample(1:nrow(yldmerge), size = 20)) {
    for(jj in 1:length(mannms)) {  # jj = 2
      a <- yldmerge[ii, ]
      b <- sapply(crind[[jj]], function(x) {
        crml[[x]][mid == a$mid, ][[ncol(crml[[x]])]]
      })
      #if(a[[paste0("YLD", mannms[jj])]] != round(mean(b))) {
      if(a[[paste0("YLD", mannms[jj])]] != round(max(b))) {
        print(paste("Yield mismatch at", ii, mannms[jj]))
      }
    }
  }

@@

soy_ylds <- fread(fp(p_yield, "DSSAT/soy_ylds.csv"))
# soy_ylds[, year := as.numeric(substr(PDAT, 1, 4))]
soy_ylds[, YR := 1:31, by = .(XNAME, TRNO)]  # unique year
soy_ylds[, c("RUNNO", "PROF", "PDAT", "MDAT", "PRCM") := NULL]
setnames(soy_ylds, "SOIL_ID...", "SOL")

# calculate mean and max as YLDS and YLDR
soy_ylds2 <- soy_ylds[,  round(mean(HWAH)), by = .(XNAME, x, y, WTH, SOL, YR)]
setnames(soy_ylds2, "V1", "YLDS")
soy_ylds2[, YLDR := soy_ylds[,  round(max(HWAH)), by = .(XNAME, YR)][, V1]]
soy_ylds2[, XNAME := NULL]

ylds$maize[, c("ID", "Zone", "zoneid") := NULL]
soy_ylds2 <- cbind(c("mid" = 1:nrow(soy_ylds2)), soy_ylds2)
setnames(soy_ylds2, "V1", "mid")

ylds <- list(ylds$maize, soy_ylds2)
names(ylds) <- c("maize", "soybean")

@@

## Soil
### Profile data
Note: Using an updated version of `read_dssat_prof.R` from the rCropMod library. It needs to be rebuilt into that library.  Here it is being sourced before runnign the next section. 
```{r, eval = FALSE}
pts <- IDr_ss[, .(x, y)]
# pts <- cbind.data.frame("x" = 27.4738917, "y" = -16.2783358)  # in Zambia
coordinates(pts) <- ~x+y
soilid <- extract(soilgrid, pts)  # grid id corresponding to point
profs <- soilshare[CELL5M %in% soilid]  # profiles at that point
profdat <- read_fullSOL_prof(solfile = full_path(p_soil, "WI.SOL"), 
                             profiles = profs$SoilProfile) 
solvars <- unique(gsub("[[:digit:]]", "", names(profdat)))
lnms <- names(profdat)

# max depth in profile
mdepth <- rowMaxs(as.matrix(profdat[, .SD, .SDcols = grep("SLB", lnms)]), 
                  na.rm = TRUE)

# use depths as weights to calculate mean values of soil parameters to 1 m
# but need to fill to stop at 1 m, and also correct for soils that don't get to 
# 1 m.  Assume values for deepest horizon where deepest < 1 m are the values to 
# 1 m. Truncate horizon that spans 1 m at 1 m, makes this deepest horizon in 
# calc. Horizon depths become weights for weighted mean. 
tst <- profdat[, .SD , .SDcols = grep("SLB", lnms)]
dtest <- function(x) {
  o <- x
  o[is.na(o)] <- -10000
  if(all(o < 100)) {
    o[which.max(o)] <- 100
  } else if(any(o > 100)) {
    o[which(o > 100)[1]] <- 100
  }
  o[o < 0] <- NA
  o
}
depvec <- t(apply(as.matrix(tst), 1, dtest))  # fix depths

@@

# calculate mean values to 1 m depth  
solvarsr <- solvars[solvars %in% 
                     c("SBDM", "SLOC", "SLCL", "SLSI", "SLHW")]
solvals <- do.call(cbind, lapply(solvarsr, function(i) {  # i <- solvarsr[2]
  mat <- as.matrix(profdat[, .SD , .SDcols = grep(i, lnms)])
  oval <- sapply(1:nrow(mat), function(j) {  # j <- 6
    ind <- which(!is.na(depvec[j, ]))
    matv <- mat[j, ind]
    matv <- infill(x = matv, ind)  # fill in NA values within profiles
    w <- depvec[j, ind] - c(0, depvec[j, ind[-length(ind)]])  # depth weights
    # weighted.mean(matv, depvec[j, ind], na.rm = TRUE)  # this was wrong
    weighted.mean(matv, w, na.rm = TRUE)
  })
  oval
}))
colnames(solvals) <- solvarsr

# finalize
sol1m <- cbind("prof" = profdat[, prof], data.table(solvals))
# unique(ylds$soybean$SOL) %in% sol1m$prof
```

### ISRIC grids
```{r, eval = FALSE}
p_spreds <- fp(p_root, "external/input_devel/soil")
mgrid <- raster("agroEcoTradeoff/external/data/ZA/ZA-mask.tif")
ext <- bbox(mgrid)[1:4]  # extent   
crpshp <- as(extent(mgrid), "SpatialPolygons")
crpshp@proj4string <- mgrid@crs

depths <- c(0, 5, 15, 30, 60, 100) 
thickness <- depths[-1] - depths[-length(depths)]

p_soc <- fp(p_sgrids, "carbon/SoilGrids1km")
fnms <- dir(p_soc, pattern = "sd[1-5]", full.names = TRUE)
socr <- stack(lapply(fnms, function(x) {
  r <- raster(x)
  r <- crop(r, spTransform(crpshp, r@crs)) / 10 # crop, convert permille to %
}))
# plot(socr[[1]])

# wmu <- function(x, na.rm = na.rm) weighted.mean(x, w = thickness, na.rm = na.rm) 
# soc1m <- calc(socr, fun = wmu, na.rm = TRUE)
soc1m <- weighted.mean(socr, w = thickness)
soc1m <- writeRaster(soc1m, filename = fp(p_spreds, "soctemp.tif"), 
                     overwrite = TRUE)
# soc1m <- raster(fp(p_spreds, "soctemp.tif"))
# weighted.mean(socr[1], thickness) == soc1m[1]

onm <- fp(p_spreds, "soctempalb.tif")
gdalwarp(srcfile = soc1m@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
soc1m <- raster(onm)

# clay grids
p_cl <- fp(p_sgrids, "clay/SoilGrids1km")
fnms <- dir(p_cl, pattern = "sd[1-5]", full.names = TRUE)
sclr <- stack(lapply(fnms, function(x) {
  r <- raster(x)
  r <- crop(r, spTransform(crpshp, r@crs))
  # r[r > 100] <- 100
  r
}))
scl1m <- weighted.mean(sclr, w = thickness)
scl1m <- writeRaster(scl1m, filename = fp(p_spreds, "claytemp.tif"), 
                     overwrite = TRUE)
# scl1m <- raster(fp(p_spreds, "claytemp.tif"))
# weighted.mean(sclr[2], thickness) == scl1m[2]
onm <- fp(p_spreds, "claytempalb.tif")
gdalwarp(srcfile = scl1m@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
scl1m <- raster(onm)

# ph
p_ph <- fp(p_sgrids, "pH/SoilGrids1km")
fnms <- dir(p_ph, pattern = "sd[1-5]", full.names = TRUE)
sphr <- stack(lapply(fnms, function(x) {
  r <- raster(x)
  r <- crop(r, spTransform(crpshp, r@crs)) / 10
  r
}))
sph1m <- weighted.mean(sphr, w = thickness)
sph1m <- writeRaster(sph1m, filename = fp(p_spreds, "phtemp.tif"), 
                     overwrite = TRUE)
# sph1m <- raster(fp(p_spreds, "phtemp.tif"))
# weighted.mean(sphr[2], thickness) == sph1m[2]

onm <- fp(p_spreds, "phtempalb.tif")
gdalwarp(srcfile = sph1m@file@name, t_srs = projection(mgrid), 
         dstfile = onm, r = "average", ot = "Float32", te = ext, 
         srcnodata = -32768, dstnodata = -99, tr = c(1000, 1000), 
         of = "GTiff", multi = TRUE, verbose = FALSE, overwrite = TRUE)
sph1m <- raster(onm)
# plot(sph1m)

soilgrids <- brick(stack(list("SLOC" = soc1m, "SLCL" = scl1m, "SLHW" = sph1m)))
soilgrids <- writeRaster(soilgrids, filename = fp(p_spreds, "soilgrids.tif"), 
                         overwrite = TRUE)
# plot(soilgrids)
```

@@

# variabes to derive: MAP, N raindays, rainfall depth, n rain days, 
# n dry days in Jan and Feb, mean max temp, N days > 30, annual mean temp, 
# GDD > 30, average daily SRAD
# set up indices for selecting dates
ystrt <- which(mos == 7 & dys == 15) 
ystrt <- ystrt[-length(ystrt)]
yend <- which(mos == 7 & dys == 14)[-1]
# set up seasonal index
sind <- rep(0, length(dts))  # whole year spanning growing season
for(i in 1:length(ystrt)) sind[ystrt[i]:yend[i]] <- i
sindu <- unique(sind)[-1]
janind <- rep(0, length(dts))  # january of each year
for(i in sindu) janind[which(sind == i & mos == 1)] <- i
febind <- rep(0, length(dts))  # feb of each year
for(i in sindu) febind[which(sind == i & mos == 2)] <- i
jfmind <- rep(0, length(dts))  # djf of each growing season
for(i in sindu) jfmind[which(sind == i & mos %in% c(1, 2, 3))] <- i
ndjfind <- rep(0, length(dts))  # djf of each growing season
for(i in sindu) ndjfind[which(sind == i & mos %in% c(1, 2, 11, 12))] <- i
octind <- rep(0, length(dts))  # feb of each year
for(i in sindu) octind[which(sind == i & mos == 10)] <- i

@@

# JFM season rainfall
jfmtot <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(jfmind == sindu2[i])
  calc(prec[[ind]], sum)
})
writeRaster(jfmtot, filename = fp(p_clim, "jfmtot.tif"), overwrite = TRUE)

@@

# growing degree day and killing degree days
kdd <- stack(kdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], kdd))
writeRaster(kdd, filename = fp(p_clim, "kdd.tif"), overwrite = TRUE)
gdd <- stack(gdy <- foreach(i = 1:length(tmu)) %dopar% calc(tmu[[i]], gdd830))
writeRaster(gdd, filename = fp(p_clim, "gdd.tif"), overwrite = TRUE)

# tmax greater than 30, Jan and Feb
nkilljan <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(janind == sindu2[i])
  calc(tmax[[ind]] - 273.15, ftmax30)
})
writeRaster(nkilljan, filename = fp(p_clim, "nkilljan.tif"), overwrite = TRUE)
nkillfeb <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(febind == sindu2[i])
  calc(tmax[[ind]] - 273.15, ftmax30)
})
writeRaster(nkillfeb, filename = fp(p_clim, "nkillfeb.tif"), overwrite = TRUE)

# mean tmax in NDJF
tmax_gs <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(ndjfind == sindu2[i])
  calc(tmax[[ind]] - 273.15, mean)
})
writeRaster(tmax_gs, filename = fp(p_clim, "tmax_gs.tif"), overwrite = TRUE)

# mean tmax whole year
tmax_ano <- stack(foreach(i = 1:length(sindu2)) %dopar% { #  i <- 1
  ind <- which(sind == sindu2[i])
  calc(tmax[[ind]] - 273.15, mean)
})
writeRaster(tmax_ano, filename = fp(p_clim, "tmax_ano.tif"), overwrite = TRUE)
